Project 4

### 2018-11-20 Project4/report/Project4.tex
Finished

\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{epsfig,floatflt}
\usepackage{url}
\usepackage{subfig}
\setcitestyle{square}

@@

\begin{abstract}

\end{abstract}
\keywords{computational science: Ising model  --- methods: Metropolis, Monte Carlo}

\section{Introduction}
\label{sec:introduction}
To study the phase transitions of a magnetic system we consider the Ising model for a two-dimensional system with a finite lattice size. The particles (spins) that make up the crystal (lattice) can take two values representing spin up or down. The energy of the lattice can then, by the Ising model, be found by
%
\begin{gather}\label{eq:energyIsing}
    E = -J\sum_{< kl >}^N s_k s_l,
\end{gather}
%
when there is no external magnetic field applied to the system. The coupling constant J will be set to 1, which means we are studying a ferromagnetic interaction between the spins that make up the lattice. Since the coupling constant is dependent upon the material we study, setting this to 1 allows for a general model. For studying the phase transition we use a Markov chain Monte Carlo method (MCMC), and the probability of finding the system in a given state $i$ is
%
\begin{align*}
    P(E_i) = \frac{e^{-\beta E_i}}{Z},
\end{align*}
%
where Z is the partition function $Z = \sum_i e^{-\beta E_i}$, which is impossible to calculate unless we know all the states. To eliminate the partition function we will use the Metropolis algorithm, which eliminates the partition function. When applying the algorithm we will consider periodic boundary conditions of the lattice, which will allow us to simulate an infinitely large lattice, and we do not have to consider what happens on the boundary the lattice (the ends of the material). Increasing the size of the lattice will further improve the approximation of a large scale lattice. The goal of the MCMC simulation is to observe the energy and the magnetic moment as time (MC cycles) increase, we expect the system to stabilize in the most probable state.

@@

The two-dimensional binary Ising model has analytical expectation values. The partition function
%
\begin{equation*}
    z = \sum_i \Omega(E_i) e^{\beta E_i},
\end{equation*}
%
where $\Omega(E_i)$ is the degeneracy of energy $E_i$ and $\beta = 1/k_BT$. From Table \ref{tab:energies} we can see that there are 2 energies that are non-zero: $\pm 8J$, which both has a degeneracy of 2. The energy equal to zero has a degeneracy of 12. Expanding the sum of the partition function we get 
%
\begin{align*}
    z &= 2e^{8J\beta} + 12 + 2e^{-8J\beta} \\
    &= 12 + 2\left(e^{8J\beta} + e^{-8J\beta}\right),
\end{align*}
%
where we have inserted $e^0 = 1$. We know that $\cosh(x) = \frac{1}{2}\left( e^{-x} + e^x\right)$, inserting this gives
%
\begin{gather}\label{eq:z}
    Z = 12 + 4\cosh(8J\beta).
\end{gather}
%
The mean value of the energy is given by
%
\begin{align*}
    \langle E \rangle = -\frac{\partial }{\partial \beta} \ln Z = -\frac{1}{Z}\frac{\partial Z}{\partial \beta},
\end{align*}
%
which is found by inserting the partition function.
%
\begin{align*}
    \langle E \rangle &= -\frac{1}{Z}\frac{\partial}{\partial \beta} \left(12 + 4 \cosh\left(8J\beta\right)\right)\\
    &= -\frac{4}{Z}\left(8J\sinh\left(8J\beta\right)\right),
\end{align*}
pulling a factor 4 from the partition function and inserting we are left with a analytical solution to the mean energy of the lattice
%
\begin{gather*}
    \langle E \rangle = -\frac{8J\sinh\left(8J\beta\right)}{3 + \cosh(8J\beta}.
\end{gather*}
%
The heat capacity can be found by
%
\begin{align*}
    C_V &= \frac{1}{kT^2}\frac{\partial^2}{\partial \beta^2} \ln Z = \frac{1}{kT^2}\frac{\partial}{\partial \beta}\left(\frac{\partial}{\partial \beta} \ln Z\right),
\end{align*}
%
where we recognize the partial derivative in the parenthesis as the negative mean energy. 
%
\begin{align*}
    C_V = \frac{1}{kT^2}&\left(\frac{64J^2\cosh\left(8J\beta\right)\left(3 + \cosh\left(8J\beta\right)\right)}{\left(3 + \cosh\left(8J\beta\right)\right)^2}\right \\
    &-\left \frac{64J^2\sinh^2\left(8J\beta\right)}{\left(3 + \cosh\left(8J\beta\right)\right)^2}\right),
\end{align*}
%
by inserting for the mean energy and using the product rule. Further simplifications gives the mean heat capacity
%
\begin{gather*}
    C_V = \frac{64J^2}{k_B T^2} \left(\frac{\cosh(8\beta J)(3+\cosh(8 \beta J)) - \sinh^2(8\beta J)}{\left(3 + \cosh(8\beta J)\right)^2}\right)
\end{gather*}
The susceptibility is given as
%
\begin{equation*}
    \chi = \frac{1}{kT}\left(\langle M^2 \rangle - \langle M \rangle^2\right).
\end{equation*}
%
To find the susceptibility of the system we consider the mean magnetic moment
%
\begin{equation*}
    \langle M \rangle = \frac{1}{Z} \sum_i |M_i| e^{-\beta E_i},
\end{equation*}
where we find the magnetic moment for a given energy in Table \ref{tab:energies}. So the mean magnetic moment is
%
\begin{equation*}
    \langle M \rangle = \frac{1}{Z}\left(4e^{8J\beta} + 4\left(2e^{0}\right) + 4\left(-2e^0\right) + -4e^{8J\beta} \right) = 0.
\end{equation*}
%
Further we calculate the mean of the  magnetic moment squared
%
\begin{equation*}
    \langle M^2 \rangle = \frac{1}{Z}\sum_i M_i^2 e^{-\beta E_i},
\end{equation*}
and again referring to Table \ref{tab:energies} we get
%
\begin{align*}
    \langle M^2 \rangle &= \frac{1}{Z}\left(16\left(2e^{8J\beta}\right) + 16\left(2e^0\right) \right) \\
    &= \frac{8\left(e^{8J\beta} + 1\right)}{3 + \cosh\left(8J\beta\right)}.
\end{align*}
So the susceptibility of the system is given as
%
\begin{equation*}
    \chi = \frac{1}{kT^2}\langle M^2 \rangle = \frac{1}{kT^2}\frac{8\left(e^{8J\beta} + 1\right)}{3 + \cosh\left(8J\beta\right)}
\end{equation*}
%
Further we find the mean of the absolute value of the magnetic moment by

@@

\subsection{Algorithm}
For a MC cycle we initialize the lattice by setting all the spins in either a random or an ordered state, and then calculating the energy and the magnetic moment. We then iterate over the lattice at random positions (i.e. we check $L^2$ random spins), for each of these positions we calculate the change in energy $\Delta E$. Consider the ratio of the probability of the new energy $P\left(E_{new}\right)$ and the previous energy $P\left(E_{prev}\right)$:
%
\begin{align*}
    \frac{P\left(E_{new}\right)}{P\left(E_{prev}\right)} = \frac{e^{-\beta E_{new}}/Z}{e^{-\beta E_{prev}}/Z} = e^{-\beta\left(E_{new} - E_{prev}\right)} = e^{-\beta\Delta E},
\end{align*}
%
where Z is the the partition function. We now perform the Metropolis test for a given $\Delta E$;
%
\begin{equation*}
    r \leq e^{-\beta \Delta E}, \qquad r \in [0,1]
\end{equation*}
%
i.e. if the ratio is larger than or equal to a random number $r \in [0,1]$ we accept a new configuration, and flip the spin.  We are looking at the change in energy if we were to flip a random spin; if the change of energy is large, the likelihood of this spin being flipped is small, so we reject this solution, on the other hand, if the probability is large, we accept the configuration. Since we expect the system to move to more likely state, we are looking for states that does not increase the systems energy. A large change in energy means we are looking at a spin that has half or more of its surrounding spins in the same configuration, and the system seems to locally move towards a steady state, therefore the local system is not changed. For each such Metropolis test that succeeds we add the flip of spin as a contribution to the systems energy and magnetic moment. When the calculations over the lattice is complete we conclude a single MC cycle and store the final value for the energy and magnetic moment.

PBC

\section{Results}
\label{sec:results}

\begin{deluxetable}{lccccc}
\tablewidth{0pt}
\tablecaption{\label{tab:analytical}}
\tablecomments{Comparison of analytical and numerical results for various values.}
\tablecolumns{4}
\tablehead{ $\langle E \rangle$ & $\langle |M| \rangle$ & $C_V$ & $\chi$}
\startdata
-7.983 & 3.994 & 0.128 & 15.973 
\enddata
\end{deluxetable}

\begin{deluxetable}{lccccc}
\tablewidth{0pt}
\tablecaption{\label{tab:results}}
\tablecomments{Comparison of analytical and numerical results for various MC cycles. Lattice size is 2x2, temperature 1 kT/J and initial spin randomly set.}
\tablecolumns{5}
\tablehead{MC cycles & $\langle E \rangle$ & $\langle |M| \rangle$ & $C_V$ & $\chi$}
\startdata
1e3 & -7.976 & 3.992 & 0.191 & 0.023 \\
5e3 & -7.984 & 3.994 & 0.127 & 0.027 \\
1e4 & -7.968 & 3.989 & 0.254 & 15.928 \\
1e5 & -7.984 & 3.994 & 0.122 & 15.403 \\
1e6 & -7.984 & 3.994 & 0.127 & 15.973 \\
Analytical & -7.983 & 3.994 & 0.128 & 15.973 
\enddata
\end{deluxetable}

\begin{figure}[h]
\subfloat[Energy pr. spin.]{
  \includegraphics[width=9cm]{EofMCC-GS-T1-L20-1e5.pdf}
  \label{fig:L20-GS-E-T1}}\\
\subfloat[Absolute value of the magnetic moment pr. spin.]{
  \includegraphics[width=9cm]{MofMCC-GS-T1-L20-1e5.pdf}
  \label{fig:evaluation:L20-GS-M-T1}}
\caption{Time evolution (MC cycles) of energy and magnetic moment of an ordered initial state with temperature 1 kT/J. $10^5$ Monte Carlo cycles.}
\end{figure}

\begin{figure}[h]
\subfloat[Energy pr. spin.]{%
  \includegraphics[width=8.65cm]{EofMCC-GS-T2_4-L20-1e5.pdf}
  \label{fig:L20-GS-E-T2.4}}\\
\subfloat[Absolute value of the magnetic moment pr. spin.]{
  \includegraphics[width=8.65cm]{MofMCC-GS-T2_4-L20-1e5.pdf}
  \label{fig:L20-GS-M-T2.4}}
\caption{Time evolution of energy and magnetic moment of an ordered initial state with temperature 2.4 kT/J. $10^5$ Monte Carlo cycles.}
\end{figure}

\begin{figure}[h]
\subfloat[Energy pr. spin.]{%
  \includegraphics[width=8.65cm]{EofMCC-T1-L20-1e6.pdf}%
  \label{fig:L20-E-T1}%
}\\
\subfloat[Absolute value of the magnetic moment pr. spin.]{%
  \includegraphics[width=8.65cm]{MofMCC-T1-L20-1e6.pdf}%
  \label{fig:L20-M-T1}%
}
\caption{Time evolution of energy and magnetic moment of an random initial state with temperature 1 kT/J. $10^6$ Monte Carlo cycles.}
\end{figure}

\begin{figure}[h]
\subfloat[Energy pr. spin.]{
  \includegraphics[width=8.65cm]{EofMCC-T2_4-L20-1e6.pdf}
  \label{fig:evaluation:revenue}}\\
\subfloat[Absolute value of the magnetic moment pr. spin.]{
  \includegraphics[width=8.65cm]{MofMCC-T2_4-L20-1e6.pdf}
  \label{fig:evaluation:avgPrice}}
\caption{Time evolution (MC cycles) of energy and magnetic moment of an random initial state with temperature 2.4 kT/J. $10^6$ Monte Carlo cycles.}
\end{figure}

\begin{figure}[h]
\mbox{\epsfig{figure=accepts_of_T-T1T2_4-L20.pdf,width=\linewidth,clip=}}
\caption{Logarithm of the accepted states as a function of the temperature for different number of Monte Carlo cycles. The left four dots corresponds to $T=1$ kT/J, and the right four dots corresponds to $T=2.4$ kT/J. The initial spins are randomly organized.}
\label{fig:logAccepts(t)}
\end{figure}

\begin{figure}[h]
\mbox{\epsfig{figure=accepts-T1T2_4-L20-1e5.pdf,width=\linewidth,clip=}}
\caption{Logarithmic plot of accepted states as a function of Monte Carlo cycles. The initial spins are randomly organized.}
\label{fig:logAccepts}
\end{figure}

\begin{figure}[h]
\mbox{\epsfig{figure=E-T2-23.pdf,width=\linewidth,clip=}}
\caption{Mean energy pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
\label{fig:E-T2-23}
\end{figure}

\begin{figure}[h]
\mbox{\epsfig{figure=M-T2-23.pdf,width=\linewidth,clip=}}
\caption{Mean absolute value of magnetic moment pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
\label{fig:M-T2-23}
\end{figure}

\begin{figure}[h]
\mbox{\epsfig{figure=CV-T2-23.pdf,width=\linewidth,clip=}}
\caption{Heat capacity pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
\label{fig:CV-T2-23}
\end{figure}

\begin{figure}[h]
\mbox{\epsfig{figure=X-T2-23.pdf,width=\linewidth,clip=}}
\caption{Susceptibility pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.03, making it ten steps.}
\label{fig:chi-T2-23}
\end{figure}

\begin{figure}[h]
\mbox{\epsfig{figure=E-T22-24.pdf,width=\linewidth,clip=}}
\caption{Mean energy pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
\label{fig:E-T22-23}
\end{figure}

\begin{figure}[h]
\mbox{\epsfig{figure=M-T22-24.pdf,width=\linewidth,clip=}}
\caption{Mean abslute value of the magnetic moment pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
\label{fig:M-T22-23}
\end{figure}

\begin{figure}[h]
\mbox{\epsfig{figure=CV-T22-24.pdf,width=\linewidth,clip=}}
\caption{Heat capacity pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
\label{fig:CV-T22-24}
\end{figure}

\begin{figure}[h]
\mbox{\epsfig{figure=X-T22-24.pdf,width=\linewidth,clip=}}
\caption{Susceptibility pr. spin for various amounts of MC cycles. Initial spin state is random. Temperature varies with a step size of 0.01, making it twenty steps.}
\label{fig:chi-T22-23}
\end{figure}

\begin{figure}[h]
\mbox{\epsfig{figure=FittedTc.pdf,width=\linewidth,clip=}}
\caption{Linear regression used on the critical temperatures at points of 1/L for $L = 40$, $60$, $80$ and $100$ gives the function $y(x) = 3.45x + 2.22$.}
\label{fig:figure_label}
\end{figure}

\begin{figure}[h]
\mbox{\epsfig{figure=PE-T1-L20-1e6.pdf,width=\linewidth,clip=}}
\caption{Histogram of the probability as a function of energy. Temperature is 1 kT/J. Initial spin states are random.}
\label{fig:figure_label}
\end{figure}

\begin{figure}[h]
\mbox{\epsfig{figure=PE-T2_4-L20-1e6.pdf,width=\linewidth,clip=}}
\caption{Histogram of the probability as a function of energy. Temperature is 2.4 kT/J. Initial spin states are random.}
\label{fig:figure_label}
\end{figure}

\begin{deluxetable}{cccc}
\tablewidth{0pt}
\tablecaption{\label{tab:runtimes40}}
\tablecomments{Run times for a 40x40 lattice for $10^5$ Monte Carlo cycles. Temperature set to 1 kT/J and the initial state for the spins are random.}
\tablecolumns{4}
\tablehead{Run & 1 core & 4 core & Difference}
\startdata
1 & 13.910 & 4.448 & 9.461 \\
2 & 14.348 & 4.530 & 9.817 \\
3 & 14.407 & 4.636 & 9.770 \\
Average & 14.221 & 4.538 & 9.683
\enddata
\end{deluxetable}

\begin{deluxetable}{cccc}
\tablewidth{0pt}
\tablecaption{\label{tab:runtimes100}}
\tablecomments{Run times for a 100x100 lattice for $10^5$ Monte Carlo cycles. Temperature set to 1 kT/J and the initial state for the spins are random.}
\tablecolumns{5}
\tablehead{Run & 1 core & 4 core & Difference}
\startdata
1 & 91.137 & 34.434 & 62.702 \\
2 & 83.662 & 30.455 & 53.207 \\
3 & 86.905 & 28.534 & 58.371 \\
Average & 87.235 & 31.141 & 58.093
\enddata
\end{deluxetable}

\section{Discussion}
\label{sec:discussion}

\section{Conclusions}
\label{sec:conclusions}

%\begin{figure}[h]
%
%\mbox{\epsfig{figure=filename.eps,width=\linewidth,clip=}}
%
%\caption{Description of figure -- explain all elements, but do not
%draw conclusions here.}
%\label{fig:figure_label}
%\end{figure}

>>>>

\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{epsfig,floatflt}
\usepackage{url}
\usepackage[caption=false]{subfig}
\usepackage{url}
\setcitestyle{square}

@@

\begin{abstract}
In this study of phase transformations of a magnetic system we will estimate the critical temperature using the Ising model. To study the system of time we will use the Metropolis algorithm, which is a Markov chain-Monte Carlo method. Since we do not know all the possible states of the system beforehand, the Metropolis algorithm is beneficial to use, since it eliminates the partition function. The study will show that for a cold system of temperature 1 kT/J the system rarely has states of higher energies, when increasing the temperature to 2.4 kT/J the total energy of the system is increased, and we see that the probability for the system to be in a higher energy state is increased. When we look at how the system evolves as a function of temperature we clearly see that there is a phase transformation at a temperature close to the critical temperature. When estimating the critical temperature we find this to be $T_C = 2.256$ kT/J, which is fairly close to the analytical value found by Lars Onsager \cite{bib:onsager}.

\end{abstract}
\keywords{computational science: Ising model  --- methods: Metropolis algorithm, Markov chain-Monte Carlo}

\section{Introduction}
\label{sec:introduction}
To study the phase transitions of a magnetic system we consider the Ising model for a two-dimensional system with a finite lattice size. The particles (spins) that make up the crystal (lattice) can take two values representing spin up or down. The energy of the lattice can then, by the Ising model, be found by
%
\begin{gather}\label{eq:energyIsing}
    E = -J\sum_{< kl >}^N s_k s_l,
\end{gather}
%
when there is no external magnetic field applied to the system. The coupling constant J will be set to 1, which means we are studying a ferromagnetic interaction between the spins that make up the lattice. Since the coupling constant is dependent upon the material we study, setting this to 1 allows for a general model. For studying the phase transition we use a Markov chain Monte Carlo method (MCMC), and the probability of finding the system in a given state $i$ is
%
\begin{align}\label{eq:P_i}
    P_i = P(E_i) = \frac{e^{-\beta E_i}}{Z},
\end{align}
%
where Z is the partition function $Z = \sum_i e^{-\beta E_i}$, which is impossible to calculate unless we know all the states. To eliminate the partition function we will use the Metropolis algorithm, which eliminates the partition function. When applying the algorithm we will consider periodic boundary conditions of the lattice, which will allow us to simulate an infinitely large lattice, and we do not have to consider what happens on the boundary the lattice (the ends of the material). Increasing the size of the lattice will further improve the approximation of a large scale lattice. The goal of the MCMC simulation is to observe the energy and the magnetic moment as time (MC cycles) increase, we expect the system to stabilize in the most probable state.

@@

The two-dimensional binary Ising model has analytical expectation values. The partition function
%
\begin{equation*}
    z = \sum_i \Omega(E_i) e^{\beta E_i},
\end{equation*}
%
where $\Omega(E_i)$ is the degeneracy of energy $E_i$ and $\beta = 1/k_BT$. From Table \ref{tab:energies} we can see that there are 2 energies that are non-zero: $\pm 8J$, which both has a degeneracy of 2. The energy equal to zero has a degeneracy of 12. Expanding the sum of the partition function we get 
%
\begin{align*}
    z &= 2e^{8J\beta} + 12 + 2e^{-8J\beta} \\
    &= 12 + 2\left(e^{8J\beta} + e^{-8J\beta}\right),
\end{align*}
%
where we have inserted $e^0 = 1$. We know that $\cosh(x) = \frac{1}{2}\left( e^{-x} + e^x\right)$, inserting this gives
%
\begin{gather}\label{eq:z}
    Z = 12 + 4\cosh(8J\beta).
\end{gather}
%
The mean value of the energy is given by
%
\begin{align*}
    \langle E \rangle = -\frac{\partial }{\partial \beta} \ln Z = -\frac{1}{Z}\frac{\partial Z}{\partial \beta},
\end{align*}
%
which is found by inserting the partition function.
%
\begin{align*}
    \langle E \rangle &= -\frac{1}{Z}\frac{\partial}{\partial \beta} \left(12 + 4 \cosh\left(8J\beta\right)\right)\\
    &= -\frac{4}{Z}\left(8J\sinh\left(8J\beta\right)\right),
\end{align*}
pulling a factor 4 from the partition function and inserting we are left with a analytical solution to the mean energy of the lattice
%
\begin{gather*}
    \langle E \rangle = -\frac{8J\sinh\left(8J\beta\right)}{3 + \cosh(8J\beta}.
\end{gather*}
%
The heat capacity can be found by
%
\begin{align*}
    C_V &= \frac{1}{kT^2}\frac{\partial^2}{\partial \beta^2} \ln Z = \frac{1}{kT^2}\frac{\partial}{\partial \beta}\left(\frac{\partial}{\partial \beta} \ln Z\right),
\end{align*}
%
where we recognize the partial derivative in the parenthesis as the negative mean energy. 
%
\begin{align*}
    C_V = \frac{1}{kT^2}\bigg(&\frac{64J^2\cosh\left(8J\beta\right)\left(3 + \cosh\left(8J\beta\right)\right)}{\left(3 + \cosh\left(8J\beta\right)\right)^2} \\ &-\frac{64J^2\sinh^2\left(8J\beta\right)}{\left(3 + \cosh\left(8J\beta\right)\right)^2}\bigg),
\end{align*}
%
by inserting for the mean energy and using the product rule. Further simplifications gives the mean heat capacity
%
\begin{gather*}
    C_V = \frac{64J^2}{k_B T^2} \left(\frac{\cosh(8\beta J)(3+\cosh(8 \beta J)) - \sinh^2(8\beta J)}{\left(3 + \cosh(8\beta J)\right)^2}\right)
\end{gather*}
The susceptibility is given as
%
\begin{equation*}
    \chi = \frac{1}{kT}\left(\langle M^2 \rangle - \langle M \rangle^2\right).
\end{equation*}
%
To find the susceptibility of the system we consider the mean magnetic moment
%
\begin{equation*}
    \langle M \rangle = \frac{1}{Z} \sum_i |M_i| e^{-\beta E_i},
\end{equation*}
where we find the magnetic moment for a given energy in Table \ref{tab:energies}. So the mean magnetic moment is
%
\begin{equation*}
    \langle M \rangle = \frac{1}{Z}\left(4e^{8J\beta} + 4\left(2e^{0}\right) + 4\left(-2e^0\right) + -4e^{8J\beta} \right) = 0.
\end{equation*}
%
Further we calculate the mean of the  magnetic moment squared
%
\begin{equation*}
    \langle M^2 \rangle = \frac{1}{Z}\sum_i M_i^2 e^{-\beta E_i},
\end{equation*}
and again referring to Table \ref{tab:energies} we get
%
\begin{align*}
    \langle M^2 \rangle &= \frac{1}{Z}\left(16\left(2e^{8J\beta}\right) + 16\left(2e^0\right) \right) \\
    &= \frac{8\left(e^{8J\beta} + 1\right)}{3 + \cosh\left(8J\beta\right)}.
\end{align*}
So the susceptibility of the system is given as
%
\begin{equation*}
    \chi = \frac{1}{kT^2}\langle M^2 \rangle = \frac{1}{kT^2}\frac{8\left(e^{8J\beta} + 1\right)}{3 + \cosh\left(8J\beta\right)}
\end{equation*}
%
Further we find the mean of the absolute value of the magnetic moment by

@@

\subsection{Implementation}
For a MC cycle we initialize the lattice by setting all the spins in either a random or an ordered state, and then calculating the energy and the magnetic moment. We then iterate over the lattice at random positions (i.e. we check $L^2$ random spins), for each of these positions we calculate the change in energy $\Delta E$. Consider the ratio of the probability of the new energy $P\left(E_{new}\right)$ and the previous energy $P\left(E_{prev}\right)$:
%
\begin{align*}
    \frac{P\left(E_{new}\right)}{P\left(E_{prev}\right)} = \frac{e^{-\beta E_{new}}/Z}{e^{-\beta E_{prev}}/Z} = e^{-\beta\left(E_{new} - E_{prev}\right)} = e^{-\beta\Delta E},
\end{align*}
%
where Z is the the partition function. We now perform the Metropolis test for a given $\Delta E$;
%
\begin{equation*}
    r \leq e^{-\beta \Delta E}, \qquad r \in [0,1]
\end{equation*}
%
i.e. if the ratio is larger than or equal to a random number $r \in [0,1]$ we accept a new configuration, and flip the spin.  We are looking at the change in energy if we were to flip a random spin; if the change of energy is large, the likelihood of this spin being flipped is small, so we reject this solution, on the other hand, if the probability is large, we accept the configuration. Since we expect the system to move to more likely state, we are looking for states that does not increase the systems energy. A large change in energy means we are looking at a spin that has half or more of its surrounding spins in the same configuration, and the system seems to locally move towards a steady state, therefore the local system is not changed. For each such Metropolis test that succeeds we add the flip of spin as a contribution to the systems energy and magnetic moment. When the calculations over the lattice is complete we conclude a single MC cycle and store the final value for the energy and magnetic moment.

\subsection{Critical temperature}
As the temperature of the system moves towards the critical temperature, the susceptibility $\chi$ and the heat capacity $C_V$ moves towards an infinite value. We can see this by
%
\begin{gather*}
    C_V \sim |T_C - T|^{-\alpha} \\
    \chi \sim |T_C - T|^{-\nu},
\end{gather*}
%
where $\alpha$ and $\nu$ are positive constants. For the results in this study we have clearer peaks for the susceptibility, so we will extract the critical temperature from these. We know that the critical tempeature scales as
%
\begin{gather*}
    T_C\left(L\right) - T_C\left(L=\infty\right) = aL^{-1/\nu},
\end{gather*}
where we set $\nu = 1$, and a is a positive constant. Solving the above equation for the critical temperature as the lattice size is infinite we get
%
\begin{gather*}
    T_C\left(L=\infty \right) \approx T_C\left(L\right) - \frac{1}{L}.
\end{gather*}

\section{Results}
\label{sec:results}
Following are the results produced in this study. The results are found in plots and tables, and the content is presented in the following subsections.

\subsection{Comparison of analytical and numerical values}
Table \ref{tab:numAnal} shows a comparison of analytical and numerical values for a 2x2 lattice.

\begin{deluxetable}{lcccc}
\tablewidth{0pt}
\tablecaption{\label{tab:numAnal}}
\tablecomments{Comparison of analytical and numerical results for various MC cycles. Lattice size is 2x2, temperature 1 kT/J and initial spin randomly set.}
\tablecolumns{5}
\tablehead{MC cycles & $\langle E \rangle$ & $\langle |M| \rangle$ & $C_V$ & $\chi$}
\startdata
1e3 & -7.984 & 3.996 & 0.127 & 0.019 \\
5e3 & -7.985 & 3.994 & 0.114 & 15.916 \\
1e4 & -7.981 & 3.994 & 0.146 & 15.769 \\
1e5 & -7.984 & 3.994 & 0.129 & 15.847 \\
1e6 & -7.984 & 3.994 & 0.127 & 15.973 \\
Analytical & -7.983 & 3.994 & 0.128 & 15.973 
\enddata
\end{deluxetable}

\subsection{Mean energy and magnetisation 20x20 lattice}
Figures \ref{fig:L20-GS-E-T1} and \ref{fig:L20-GS-M-T1} shows the energy and magnetisation of the ground state for temperature 1 kT/J, while figures \ref{fig:L20-GS-E-T2.4} and \ref{fig:L20-GS-M-T2.4} shows the same for temperature 2.4 kT/J. We see that the system stabilizes quickly. For the non-ground state plots see figures \ref{fig:L20-E-T1} and \ref{fig:L20-M-T1} for temperature 1 kT/J and figures \ref{fig:L20-E-T2.4} and \ref{fig:L20-M-T2.4} for temperature 2.4 kT/J. For the system that starts in a random spin state we see by figure \ref{fig:EM-T1_zoom} that the system stabilizes at about 250 MC cycles.
%
\begin{figure}
\subfloat[Energy pr. spin.]{
  \includegraphics[width=8.65cm]{EofMCC-GS-T1-L20-1e5.pdf}
  \label{fig:L20-GS-E-T1}} \\
\subfloat[Absolute value of the magnetic moment pr. spin.]{
  \includegraphics[width=8.65cm]{MofMCC-GS-T1-L20-1e5.pdf}
  \label{fig:L20-GS-M-T1}}
\caption{Time evolution (MC cycles) of energy and magnetic moment of an ordered initial state with temperature 1 kT/J. $10^5$ Monte Carlo cycles.}
\label{fig:EM-GS-T1}
\end{figure}
%
\begin{figure}
\subfloat[Energy pr. spin.]{%
  \includegraphics[width=8.65cm]{EofMCC-GS-T2_4-L20-1e5.pdf}
  \label{fig:L20-GS-E-T2.4}} \\
\subfloat[Absolute value of the magnetic moment pr. spin.]{
  \includegraphics[width=8.65cm]{MofMCC-GS-T2_4-L20-1e5.pdf}
  \label{fig:L20-GS-M-T2.4}}
\caption{Time evolution of energy and magnetic moment of an ordered initial state with temperature 2.4 kT/J. $10^5$ Monte Carlo cycles.}
\label{fig:EM-GS-T2.4}
\end{figure}
%
\begin{figure}
\subfloat[Energy pr. spin.]{
  \includegraphics[width=8.65cm]{EofMCC-T1-L20-1e6.pdf}
  \label{fig:L20-E-T1}} \\
\subfloat[Absolute value of the magnetic moment pr. spin.]{%
  \includegraphics[width=8.65cm]{MofMCC-T1-L20-1e6.pdf}
  \label{fig:L20-M-T1}}
\caption{Time evolution of energy and magnetic moment of an random initial state with temperature 1 kT/J. $10^6$ Monte Carlo cycles.}
\label{fig:EM-T1}
\end{figure}
%
\begin{figure}
\subfloat[Zoom of figure \ref{fig:L20-E-T1}.]{
  \includegraphics[width=8.65cm]{EofMCC-T1-L20-1e6_zoom.pdf}
  \label{fig:L20-E-T1_zoom}} \\
\subfloat[Zoom of figure \ref{fig:L20-M-T1}.]{%
  \includegraphics[width=8.65cm]{MofMCC-T1-L20-1e6_zoom.pdf}
  \label{fig:L20-M-T1_zoom}}
\caption{Zoom of previous plot to show where the energy and magnetisation stabilized.}
\label{fig:EM-T1_zoom}
\end{figure}
%
\begin{figure}
\subfloat[Energy pr. spin.]{
  \includegraphics[width=8.65cm]{EofMCC-T2_4-L20-1e6.pdf}
  \label{fig:L20-E-T2.4}} \\
\subfloat[Absolute value of the magnetic moment pr. spin.]{ 
  \includegraphics[width=8.65cm]{MofMCC-T2_4-L20-1e6.pdf}
  \label{fig:L20-M-T2.4}}
\caption{Time evolution (MC cycles) of energy and magnetic moment of an random initial state with temperature 2.4 kT/J. $10^6$ Monte Carlo cycles.}
\label{fig:EM-T2.4}
\end{figure}

\subsection{Accepted states in the Metropolis algorithm}
The accepted states as a function of temperature and MC cycles are found in figures \ref{fig:logAccepts(t)} and \ref{fig:logAccepts} respectively, both are logarithmic plots.
\begin{figure}
\mbox{\epsfig{figure=accepts_of_T-T1T2_4-L20.pdf,width=\linewidth,clip=}}
\caption{Logarithm of the accepted states as a function of the temperature for different number of Monte Carlo cycles. The left four dots corresponds to $T=1$ kT/J, and the right four dots corresponds to $T=2.4$ kT/J. The initial spins are randomly organized.}
\label{fig:logAccepts(t)}
\end{figure}
\begin{figure}
\mbox{\epsfig{figure=accepts-T1T2_4-L20-1e5.pdf,width=\linewidth,clip=}}
\caption{Logarithmic plot of accepted states as a function of Monte Carlo cycles. The initial spins are randomly organized.}
\label{fig:logAccepts}
\end{figure}

\subsection{Probability of energy}
In figures \ref{fig:histT1} and \ref{fig:histT2.4} we can see the probability distribution for a given energy for temperature 1 kT/J and 2.4 kT/J respectively. The variance of the energy $\sigma_E^2 = 9.398$ for temperature 1 kT/J and $\sigma_E^2 = 3264.7$ for temperature 2.4 kT/J.
%
\begin{figure}
\mbox{\epsfig{figure=PE-T1-L20-1e6.pdf,width=\linewidth,clip=}}
\caption{Histogram of the probability as a function of energy. Temperature is 1 kT/J. Initial spin states are random. 1e6 MC cycles.}
\label{fig:histT1}
\end{figure}
%
\begin{figure}
\mbox{\epsfig{figure=PE-T2_4-L20-1e6.pdf,width=\linewidth,clip=}}
\caption{Histogram of the probability as a function of energy. Temperature is 2.4 kT/J. Initial spin states are random. 1e6 MC cycles.}
\label{fig:histT2.4}
\end{figure}
